---
title: "Regression type Decision Tree Example"
author: "Nipunjeet Gujral"
output: 
  html_document:
  theme: flatly
  highlight: haddock
---

### General Process 
  
  >
    * Test and transform data for property of Normality
    * Validate data and optimze for complexity parameter
    * Visulaize potential relations between input variables
    * Grow Descision Tree
    * Visulaize implications of decsion tree to ipnut variables
    * Prune Decsion Tree for using derived complexity parameters
    * Visualize differnce between Pruned and Unpruned Decsion trees




| Libraries |         Purpose           |
|-----------|---------------------------|
| tidyverse | Data manipulation         |
| plotly    | Graphical visualization   |
| knitr     | Datatable visulaization   |
| ISLR      | Example data              |
| tree      | Descision Tree Production |      
| rpart     | Recursive Partitioning    |
| caret     | Data Validation           |


```{r Libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(plotly)
library(ISLR)
library(knitr)
library(tree)
library(rpart)
library(caret)
```


#### Testing Data for normality
```{r Examine Data, message=FALSE, warning=FALSE}
ISLR::Hitters$Salary %>% shapiro.test()

plotly::plot_ly(data = ISLR::Hitters,
                x = ~Salary,
                type = "histogram")
```


#### Testing log-transformed Data for normality
```{r message=FALSE, warning=FALSE}
data <- ISLR::Hitters %>% 
  dplyr::select(Hits, Years, Salary) %>%
  tidyr::drop_na() %>% 
  dplyr::mutate(Log_Salary = log(Salary))

data$Log_Salary %>% shapiro.test()

plotly::plot_ly(data = data,
                x = ~Log_Salary,
                type = "histogram")
```


#### Validating the dataset
```{r}
train_control <- caret::trainControl(method = "repeatedcv", number = 10, repeats = 3)


model = caret::train(Log_Salary ~ Years + Hits,
                     data = data, 
                     trControl = train_control, method = "rpart")
print(model)
```


#### Visualize Data 
```{r message=FALSE, warning=FALSE}
knitr::kable(data[1:10, ], caption = "Data Sample", float.environmnet = 'Sidewaystable')

```


#### Graphing relationship between `Years` and `Hits`
```{r Year vs. Hits, message=FALSE, warning=FALSE}
plot1 <- plotly::plot_ly(data = data,
                x = ~Years, y = ~Hits,
                alpha = 0.6,
                marker = list(size = 10,
                              line = list(color = "black", width = 0.5)),
                text = ~paste("Salary/1000: $", Salary, sep = ""), 
                type = "scatter") %>% 
  plotly::layout(title = "Years vs. Hits",
                 xaxis = list(zeroline = FALSE),
                 yaxis = list(zeroline = FALSE)) 
plot1
```


#### Building a Decsion Tree
```{r Build a tree, message=FALSE, warning=FALSE}
# using the tree package ####
  #Tree <- tree::tree(data = data,
  #                   formula = Log_Salary ~ Hits + Years,
  #                   split = "deviance",
  #                   method = "recursive.partition", 
  #                   model = TRUE)
  #summary(Tree)
  #plot(Tree); text(Tree)

  
# using the rpart package ####
  Tree <- rpart::rpart(data = data,
                      formula = Log_Salary ~ Years + Hits,
                      method = "anova",
                      model = TRUE)
  plot(Tree); text(Tree, all= TRUE, cex = 0.7)  
  summary(Tree)
  
```


#### Apply Tree to Data
```{r message=FALSE, warning=FALSE}
plot2 <- plotly::plot_ly(data = data,
                x = ~Years, y = ~Hits,
                alpha = 0.6,
                marker = list(size = 10,
                              line = list(color = "black", width = 0.5)),
                mode = "markers",
                text = ~paste("Salary/1000: $", Salary, sep = ""), 
                type = "scatter",
                name = "Salary") %>% 
  plotly::layout(title = "Years vs. Hits",
                 xaxis = list(zeroline = FALSE),
                 yaxis = list(zeroline = FALSE)) 
```


```{r message=FALSE, warning=FALSE}
plotly::subplot(plot1, plot2, shareX = TRUE, shareY = FALSE) %>% 
  plotly::layout(width = 1200, height = 700)
```


#### Pruning 
```{r cross validation Pruning, message=FALSE, warning=FALSE}
rpart::printcp(Tree)

Pruned_Tree <- rpart::prune(Tree, cp = 0.04446021)
plot(Pruned_Tree); text(Pruned_Tree, all = TRUE, cex = 0.7)
summary(Pruned_Tree)
```


